---
title: "Untitled"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
activity <- read.csv("activity.csv", header=TRUE)
# convert character date to POSIX date
activity$date <- as.POSIXct(strptime(activity$date, "%Y-%m-%d"),tz="")
# first convert integer time to character and pad with leading zeros...
activity$time     <- sprintf("%04d", activity$interval)#C style language, fill in leading zeros
# ...then convert to the date type
activity$time     <- as.POSIXct(activity$time, "%H%M",tz="")
head(activity)
str(activity)

```

## GitHub Documents

This is an R Markdown format used for publishing markdown documents to GitHub. When you click the **Knit** button all R code chunks are run and a markdown file (.md) suitable for publishing to GitHub is generated.

## Including Code

##You can include R code in the document as follows:

total_steps_by_date <- aggregate(list(total_steps = activity$steps),
                                 by=list(date = activity$date),
                                 FUN=sum,
                                 na.rm=TRUE)
                                 
mean(total_steps_by_date$total_steps)
median(total_steps_by_date$total_steps,na.rm=T)

#What is the average daily activity pattern?

average_steps_by_time <- aggregate(list(average_steps = activity$steps),
                                   by=list(time = activity$time,
                                           interval = activity$interval),
                                   FUN=mean,
                                   na.rm=TRUE)
average_steps_by_time[which.max(average_steps_by_time$average_steps),]   

#Imputing missing values
sum(is.na(activity[,"steps"]))
# "join" the two data frames using merge()
activity_imputed <- merge(activity,average_steps_by_time,by="interval")
# correct the NA steps with average steps for the interval
activity_imputed <- within(activity_imputed,
                           steps <- ifelse(is.na(activity_imputed$steps),
                           activity_imputed$average_steps,
                           activity_imputed$steps))
                        
total_steps_by_date_imputed <- aggregate(list(total_steps = activity_imputed$steps),
                                         by=list(date = activity_imputed$date),
                                         FUN=sum,
                                         na.rm=FALSE)
#Are there differences in activity patterns between weekdays and weekends?
# first add a character column for day of the week
activity_imputed$weekday  <- weekdays(activity_imputed$date)
# now populate a new factor column using day of the week and a simple function
activity_imputed$weekend_indicator <- as.factor(apply(activity_imputed["weekday"], 1, function(x) {
  switch(x,
         "Sunday" = "weekend",
         "Saturday" = "weekend",
         "weekday")
}))
# confirm that we have the character and factor types we expect
str(activity_imputed)
## Including Plots

# frequencies
hist(total_steps_by_date$total_steps,
     breaks=30,
     xlab="Total Steps",
     main="Total Steps Per Day",
     col="lightblue")
# desnsity
plot(density(total_steps_by_date$total_steps,
             na.rm=TRUE),
     xlab="Total Steps",
     ylab="Density",
     main="Total Steps Per Day",     
     col="purple",
     lwd=3)
par(mfrow=c(1,1))

#Daily activity pattern
plot(average_steps ~ time,
     data=average_steps_by_time,
     xlab="Time interval",
     ylab="Mean steps",
     main="Mean Steps By Time Interval",
     type="l",
     col="blue",
     lwd=2)
#Imputed values
par(mfrow=c(1,2))
# frequencies
hist(total_steps_by_date_imputed$total_steps,
     breaks=30,
     xlab="Total Steps",
     main="Total Steps Per Day",
     col="lightblue")
##Imputed values
par(mfrow=c(1,2))
# frequencies
hist(total_steps_by_date_imputed$total_steps,
     breaks=30,
     xlab="Total Steps",
     main="Total Steps Per Day",
     col="lightblue")
# desnsity
plot(density(total_steps_by_date_imputed$total_steps,
             na.rm=TRUE),
     xlab="Total Steps",
     ylab="Density",
     main="Total Steps Per Day",     
     col="purple",
     lwd=3)
par(mfrow=c(1,1))

##Now draw a panel plot using ggplot2, comparing activity patterns on weekdays and weekends.
```{r}

```

average_steps_by_time_weekend <- aggregate(list(average_steps = activity_imputed$steps),
                                           by=list(time       = activity_imputed$time.x,
                                                   daytype    = activity_imputed$weekend_indicator),
                                           FUN=mean)
library(ggplot2)
qplot(x = time,
      y = average_steps,
      geom="path",
      data = average_steps_by_time_weekend, 
      xlab="Time interval",
      ylab="Average steps",
      main="Activity Patterns\nWeekdays vs. Weekends",
      facets = daytype ~ .)
